---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# jstor

[![Travis build status](https://travis-ci.org/tklebel/jstor.svg?branch=master)](https://travis-ci.org/tklebel/jstor)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tklebel/jstor?branch=master&svg=true)](https://ci.appveyor.com/project/tklebel/jstor)
[![CRAN status](http://www.r-pkg.org/badges/version/jstor)](https://cran.r-project.org/package=jstor)

The tool [Data for Research (DfR)](http://www.jstor.org/dfr/) by JSTOR is a
valuable source for citation analysis and text mining. `jstor`
provides functions and suggests workflows for importing
datasets from DfR.

The package `jstor` provides functions and suggests workflows for importing
datasets from DfR. It was developed to deal with very large datasets which
require an agreement, but can be used with smaller ones as well.

The most important set of functions is a group of `find_*` functions:

- `find_metadata`
- `find_authors`
- `find_references`
- `find_footnotes`
- `find_fulltext`

All functions which are concerned with meta data (therefore excluding
`find_fulltext`) operate along the same lines:

1. The file is read with `xml2::read_xml()`.
2. Content of the file is extracted via XPATH or CSS-expressions.
3. The resulting data is returned in a `data.frame`.


## Installation

You can install `jstor` from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tklebel/jstor")
```

## Usage
In order to use `jstor`, you first need to load it:
```{r}
library(jstor)
```

The basic usage is simple: supply one of the `find_*`-functions with a path and
it will return a data.frame with the extracted information.
```{r}
find_metadata(jstor_example("sample_with_references.xml"))

find_authors(jstor_example("sample_with_references.xml"))
```

Further explanations, especially on how to use jstor's functions for importing
many files, can be found in the vignettes.
