---
title: Analysing n-grams with `jstor` for R
author: "Thomas Klebel"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
      toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The service [DfR](http://www.jstor.org/dfr/) by JSTOR offers several ways for
text analysis of scientific articles. 
In this vignette I will demonstrate how to analyse n-grams which DfR delivers.

Let's suppose, we are interested in the topic of "inequality" within the 
discipline of sociology. Social inequality can be considered a prime subject of
sociological inquiry. In order to gain some context on the subject, we might be
interested to analyse frequently occurring terms. DfR offers different grades of
tokenization: n-grams for 1-3 words, e.g. unigrams, bigrams and trigrams. In
case you are unfamiliar with the analysis of tokenized text, you could read
the first few paragraphs of chapters 1 and 4 in https://www.tidytextmining.com
as an introduction.

Our analysis starts at the main page of [DfR](http://www.jstor.org/dfr/).
We create a dataset^[An introduction on how to create datasets can be found
on the page of DfR: http://www.jstor.org/dfr/about/creating-datasets]
by searching for
"inequality" and selecting "sociology" as our subject. To trim down the number
of articles, we only select articles from 1997 to 2017. After
logging in/creating an account, we select unigrams and bigrams. The resulting 
`.zip`-file can be downloaded from the email which DfR sends.


Up-front, we need to load some packages. `jstor` is  available from
CRAN, so it can be installed via `install.packages("jstor")`.


```{r, message=FALSE}
library(jstor)
library(tidyverse)
library(future)

# set a lighter theme for plots
theme_set(theme_bw())

# import files in parallel
plan(multiprocess)
```


With the latest version of `jstor` we can now directly import files from a zip
archive. We only need to specify, where the zip archive is located, which
parts we want to extract, and where the resulting files should be saved. 

Before importing data, we can take a quick look at the contents of each
archive with `jst_preview_zip()`:

```{r}
jst_preview_zip("part1.zip")
```
```{r}
jst_preview_zip("part2.zip")
```

Although we could import all ngrams at this point, this would be extremely 
inefficient. It is thus best first to decide, which articles we want to analyze,
and import the corresponding ngram-files afterwards.

The following code assumes that you follow a workflow
organised around projects within RStudio (refer to
http://r4ds.had.co.nz/workflow-projects.html for further information). 

```{r, eval=FALSE}
import_spec <- jst_define_import(article = jst_get_article)

jst_import_zip("part1.zip", import_spec = import_spec, out_file = "part1")
jst_import_zip("part2.zip", import_spec = import_spec, out_file = "part2")

#> Processing files for journal_article with functions jst_get_article
#> Processing chunk 1/1
#>  Progress: ───────────────────────────────────────────────────────────── 100%
```

Since `jst_import_zip` writes the results to disk, we need to read the metadata
from the newly created file. This is made easy by `jst_re_import` which ensures,
that the data are read with the right column types.

```{r}
imported_metadata <- c("part1_journal_article_jst_get_article-1.csv",
                       "part2_journal_article_jst_get_article-1.csv") %>% 
  map_df(jst_re_import)
imported_metadata
```

# Cleaning the data
Data from DfR is inherently messy. To fix a few common issues, we can use 
`jst_augment()`:

```{r}
imported_metadata <- jst_augment(imported_metadata, quietly = TRUE)
```

For more information on common quirks with data from DfR and how to deal with
them, take a look at the `vignette("known-quirks")`.


# Exploration
Before diving into the analysis of n-grams, we might wish to take an exploratory
look at our metadata. The first thing to look at are the types of articles.
```{r}
ggplot(imported_metadata, aes(article_type)) +
  geom_bar() +
  coord_flip()
```

We can see, that the majority of articles are proper "research-articles", which
together with book-reviews and miscellaneous articles amount to ~99% of all
articles.

```{r}
imported_metadata %>% 
  count(article_type, sort = T) %>% 
  mutate(perc = scales::percent(n/sum(n)))
```

We must be cautious, however, when using this variable to distinguish articles
into categories. In this instance, we have "research-articles" which are
actually book-reviews:

```{r}
imported_metadata %>% 
  filter(article_type == "research-article" & str_detect(article_title, "Book")) %>% 
  select(file_name, article_title, pub_year)
```

For the current demonstration, we want to restrict the type of articles to
research articles, therefore we need to take steps to remove book reviews and other
miscellaneous articles: First, filter by `article_type`, then remove articles
where the title starts with "Book Review".

```{r}
research_articles <- imported_metadata %>% 
  filter(article_type == "research-article") %>% 
  filter(!str_detect(article_title, "^Book Review"))
```


## The moving wall - filtering articles by time
Since JSTOR has a [moving wall](https://support.jstor.org/hc/en-us/articles/115004879547-JSTOR-s-Moving-Wall-Archive-vs-Current-Definitions),
we should take a look at the number of articles per year in our dataset.
```{r}
research_articles %>% 
  ggplot(aes(pub_year)) +
  geom_bar() 
```

From this graph we can see an increase in research articles until 2010, after
which the number of articles
first tapers off, and then drops off sharply. For this reason we should
exclude articles at least
from 2015 onward, since the sample might get quite biased toward specific
journals.

```{r}
without_wall <- research_articles %>% 
  filter(pub_year < 2015)
```

## Flagship journals - filtering articles by journal

Since the amount of articles is still rather large for this demonstration, we
could select only a few journals. Here, we will look at articles from two
leading journals within the discipline, "American Journal
of Sociology" and "American Sociological Review". 


Since we cleaned the identifiers for journals with `jst_augment` earlier,
we can select our two flagship-journals very easily.

```{r}
flagship_journals <- without_wall %>%
  filter(journal_id %in% c("amerjsoci", "amersocirevi"))
```


# Importing bigrams
> Disclaimer: Much of the following analysis was inspired by the book
"Text Mining with R" by Julia Silge and David Robinson:
https://www.tidytextmining.com

For this demonstration we will look at bigrams to find the most common pairs of
words. Until now, we were only dealing with the metadata, therefore we need a
way to link our reduced dataset to the bigram files from DfR. The directory
structure for deliveries from DfR looks something like this:

```r
receipt-id-123456-part-001
  -- metadata
    -- journal_article_foo.xml
       .
       .
       .
  -- ngram2
    -- journal_article_foo.txt
       .
       .
       .
receipt-id-123456-part-002
  -- metadata
    -- journal_article_bar.xml
       .
       .
       .
  -- ngram2
    -- journal_article_bar.txt
       .
       .
       .
```

From this structure we can see, that the file name can serve as an identifier to
match articles and n-grams, since it is similar between metadata and
n-grams. 


To make importing a subset of ngrams more convenient, we can use
`jst_subset_ngrams`. This function returns a list of "zip-locations", which
`jst_get_ngram` can read.

```{r}
ngram_selection <- jst_subset_ngrams(c("part1.zip", "part2.zip"), "ngram2",
                                    flagship_journals)
head(ngram_selection, 2)
```



```{r, eval=FALSE}
ngram_selection <- jst_subset_ngrams(c("part1.zip", "part2.zip"), "ngram2",
                                    flagship_journals)

imported_bigrams <- ngram_selection %>% 
  furrr::future_map_dfr(jst_get_ngram)
```


```{r, echo=FALSE, include=TRUE, eval=TRUE}
# write_csv(imported_bigrams, "imported_bigrams.csv")

imported_bigrams <- data.table::fread("imported_bigrams.csv", showProgress = F)
```

From the `r dim(flagship_journals)[1]` articles in our two flagship journals we
now have `r format(dim(imported_bigrams)[1], big.mark = ",")` bigrams. The 
bigrams are calculated by
JSTOR for each article independently. In order to reduce the sample to the most 
common bigrams, we have two choices: either to include only terms which occur
*within each* article a given amount of times, or to include terms which occur 
*within all* articles a given amount of times. By only including terms which 
occur more than 5 times in each article, we can drastically reduce the number of
terms. However, we might miss some important ones: there might be terms which
do not occur repeatedly within articles, but are present in all of them.

For demonstration purposes we are a bit restrictive and include only those
terms, which occur at least three times per article. 

```{r}
top_bigrams <- imported_bigrams %>%
  filter(n >= 3)
```



## Cleaning up bigrams
When constructing n-grams, DfR uses a stop-word list, which is quite limited 
^[for more information see the 
[technical specifications](http://www.jstor.org/dfr/about/technical-specifications)
on their page]. If we would like to restrict the terms a bit further, we could
use stopwords from `tidytext`:


```{r}
library(tidytext)
bigrams_separated <- top_bigrams %>%
  separate(bigrams, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)
```

After removing the stopwords we need to consider the fact, that our bigrams
were created for each article on its own. In order to analyse them together,
we need to count the terms for all articles in combination.

```{r}
bigram_counts <- bigrams_filtered %>%
  group_by(word1, word2) %>%
  summarise(n = sum(n)) %>%
  arrange(desc(n))

bigram_counts
```

From the first few terms we can see, that there are still many terms which are
not very interesting for our analysis. The terms "american" and "sociological"
are simply part of the title of a journal we selected (American Sociological 
Review). To clean the terms up, we can employ different approaches. One is to
simply filter the terms we wish to exclude:

```{r}
bigram_counts_clean <- bigram_counts %>%
  unite(bigram, word1, word2, sep = " ") %>%
  filter(!bigram %in% c("american sociological", "sociological review",
                        "university press", "american journal",
                        "journal sociology")) %>%
  separate(bigram, c("word1", "word2"))

```

We will look at another approach after plotting our bigrams.


# Visualize relationships
When analyzing bigrams, we might want to look at the relationships between
common terms. For this we can leverage the power of
[igraph](http://igraph.org/r/) and
[ggraph](https://cran.r-project.org/web/packages/ggraph/index.html).

```{r, message=FALSE}
library(igraph)
library(ggraph)
```


First, we only keep the most common terms and then convert our `data.frame` to
an `igraph`-object. ^[If you are unfamiliar with graph theory, just take a look
at Wikipedia: [Graph Theory](https://en.wikipedia.org/wiki/Graph_theory).]


```{r}
bigram_graph <- bigram_counts_clean %>%
  filter(n > 500) %>%
  graph_from_data_frame()

bigram_graph
```

For plotting, we will use a simple plotting function, adapted from
https://www.tidytextmining.com/ngrams.html#visualizing-a-network-of-bigrams-with-ggraph.

```{r}
plot_bigrams <- function(igraph_df, seed = 2016) {
  set.seed(seed)
  
  a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

  ggraph(igraph_df, layout = "fr") +
    geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                   arrow = a, end_cap = circle(.07, 'inches')) +
    geom_node_point(color = "lightblue", size = 4) +
    geom_node_text(aes(label = name), repel = T) +
    theme_graph()
}
```

```{r, fig.width=12, fig.height=10}
plot_bigrams(bigram_graph)
```
Very obvious is a group of nodes which are not relevant to the topic of
inequality. They come from LaTeX documents and somehow made their way into the
original dataset. However, since they are more common than most of the other
terms, they are quite easy to remove. We can look at the nodes/vertices of our
graph with `V(bigram_graph)`.

```{r}
V(bigram_graph)
```

The first node, "labor", is relevant to us, but all other nodes from 2 to at 
least 40 are clearly irrelevant. We can remove them by simple subtraction:

```{r}
bigram_graph_clean <- bigram_graph - 2:40
bigram_graph_clean
```

Another apparent group is a combination of "table" or "figure" with digits. This
evidently comes from tables or figures in the papers and might suggest, that the
articles in our sample quite frequently employ quantitative methods, where
figures and tables are very common. For the analysis at hand however, we might
remove them, along with a few other irrelevant terms.


```{r}
bigram_graph_clean <- bigram_graph_clean - c("table", "model",
                                             as.character(0:5),
                                             "xd", "rh", "landscape", "00",
                                             "figure", "review", "79",
                                             "http", "www", "000", "01")
```


After cleaning up a bit, we can take a fresh look at our bigrams.
```{r, fig.width=12, fig.height=10}
plot_bigrams(bigram_graph_clean, 234)
```

The figure is still far from perfect ("eco" -> "nomic" should clearly be one
term), but we can begin to analyse our network.

The most frequent bigrams are now "labor market", "labor force", and "income
inequality", which are not very surprising given that most individuals in
capitalist societies need to supply their work in exchange for income. For this
reason, the labor market and its stratification is a prime subject of the 
sociological inquiry into inequality. A few further key dimensions of
sociological analysis are apparent from the
graph: gender, race/ethnicity, occupational and socioeconomic status. That we
find many terms to be associated with the term "social" seems quite likely
given the discipline's subject.

At least two surprising results should be pointed out. First, it is not evident
how the terms "ethnic" and "racial" are connected. They do not form a typical
term like "social capital", "middle class" or similar, nor could they be 
considered a dichotomy like "black" and "white" which are often included in
tables from regressions. From a theoretical point of view, they have slightly
different meanings but are frequently being used as synonyms. 
Second, there is a group of nodes around the term
"university": university -> chicago, university -> california,
harvard -> university, etc. At least two explanations seem plausible: either,
many books are being cited which are in some way associated with those 
universities ("The University of Chicago Press" is the largest university press
in the United States), or many researchers who publish in the two
flagship-journals we selected are affiliated with those four universities: 
Harvard, Chicago, Cambridge and California. At least partly the prominence of
university -> chicago -> press might be due to the fact, that it is the
publisher of the American Journal of Sociology, and therefore included in each
article by this journal.

Altogether, these findings will not be surprising to well-educated sociologists.
Almost all bigrams in the graph are common concepts or terms within the
discipline. Most importantly, very often those concepts and terms comprise two
single words, as in "social science", "social structure", "social networks".
Two approaches might be useful to examine, *how* these concepts are being used:

1. Analyzing trigrams. It could be the case, that many of the above concepts 
would show up in combination with more interesting terms, if we were to analyze
combinations of three words. 
2. Another approach would be to first analyze data for bigrams as above, 
determining the core concepts of a field. In a second step, one could search for
those core concepts in the raw data files and extract adjacent terms. This would
not be possible with the default deliveries from DfR however, since full-text
content is only available through a dedicated agreement.

# Comparison over time
Besides looking at the overall relationship of bigrams, we could be interested
in the development over time of specific terms. Here, we want to look at how
often "labor market" and "income inequality" appear from year to year.

For this, we need to join our bigrams with the metadata.


```{r}
time_bigrams <- top_bigrams %>% 
  left_join(flagship_journals, by = "file_name") %>% 
  select(bigrams, n, pub_year)

head(time_bigrams)
```

Again, we need to sum up the counts, but this time grouped by year:
```{r}
time_bigrams <- time_bigrams %>%
  group_by(bigrams, pub_year) %>%
  summarise(n = sum(n)) %>%
  arrange(desc(n))

time_bigrams
```

We now only keep the two terms of interest and plot them in a simple chart.

```{r, fig.width=10}
# filter the terms of interest
time_comparison <- time_bigrams %>% 
  filter(bigrams == "labor market" | bigrams == "income inequality")

ggplot(time_comparison, aes(pub_year, n, colour = bigrams)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = scales::pretty_breaks(7))
```

In this instance, the plot does not reveal trends over time-the frequency of 
the terms is fluctuating a lot but staying on a similar level. Single spikes
of term frequency for specific years (for example income inequality in 2011)
could stem from special issues being explicitly concerned with income
inequality, although a quick glance at the corresponding issues invalidates this
hypothesis.
