---
title: "Automating File Import"
author: "Thomas Klebel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Intro
```{r}
library(stringr)
```

# List files
From my knowledge, there are two ways to generate a list of all files: 
`list.files` or using `system` in conjunction with `find` on unix-like systems.

```{r, echo=FALSE}
my_dir <- system.file("extdata", package = "jstor")
```

## `list.files`

```{r}
file_names_listed <- list.files(path = my_dir, full.names = T, pattern = "*.xml")
file_names_listed
```

## `system` and `find`
```{r}
file_names <- system(paste0("cd ", my_dir, "; find . -name '*.xml' -type f"), intern = TRUE)
```

```{r}
file_names_system <- file_names %>%
  str_replace("^\\.\\/", "") %>%
  str_c(my_dir, "/", .)
file_names_system
```

In this case the two approaches give the same result.
```{r}
identical(file_names_listed, file_names_system)
```

The main difference seems to be though, that `list.files` sorts the output, 
wherase `find` does not. For a large amout of files (200,000) this makes a
difference, for smaller datasets the difference shouldn't make an impact.

# Batch import
