---
title: "Introduction to `jstor`"
author: "Thomas Klebel"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The tool [Data for Research (DfR)](http://www.jstor.org/dfr/) by JSTOR is a
valuable source for citation analysis and text mining. `jstor`
provides functions and suggests workflows for importing
datasets from DfR.

When using DfR, requests for datasets can
be made for small excerpts (max. 25,000 records) or large ones, which require an
agreement between the researcher and JSTOR. `jstor` was developed to deal with
very large datasets which require an agreement, but can be used with smaller
ones as well.

The most important set of functions is a group of `find_*` functions:

- `find_article`
- `find_authors`
- `find_references`
- `find_footnotes`
- `find_fulltext`

I will demonstrate their usage using the
[sample dataset](http://www.jstor.org/dfr/about/sample-datasets)
which is provided by JSTOR on their website.

## General Concept
All functions which are concerned with meta data (therefore excluding
`find_fulltext`) operate along the same lines:

1. The file is read with `xml2::read_xml()`.
2. Content of the file is extracted via XPATH or CSS-expressions.
3. The resulting data is returned in a `data.frame`.

The following sections showcase the different functions.


## Application
Apart from `jstor` we only need to load `dplyr` for matching records.

```{r, message=FALSE, warning=FALSE}
library(jstor)
library(dplyr)
```



### `find_article`
The basic usage of the `find_*` functions is very simple. They take only one 
argument, the path to the file to import:
```{r}
meta_data <- find_article(file_path = jstor_example("sample_with_references.xml"))
```

The resulting object is a `data.frame` with one row and 12 columns. The columns
correspond to most of the elements documented here: http://www.jstor.org/dfr/about/technical-specifications.

The columns are:

- journal_id *(chr)*: The jcode or a DOI. If both are present, the jcode
(=publisher-id) is extracted.
- basename_id *(chr)*: The filename of the original .xml-file. Can be used 
 for joining with full-texts.
- article_id *(chr)*: The article id, either a combination of digits 
 (`pub-id-type="jstor"`) or a DOI (`pub-id-type="doi"`).
- article_title *(chr)*: The title of the article.
- volume *(chr)*: The volume the article was published in.
- issue *(chr)*: The issue the article was published in.
- language *(chr)*: The language of the article.
- pub_day *(int)*: Publication day, if specified.
- pub_month *(int)*: Publication month, if specified.
- pub_year *(int)*: Year of publication.
- first_pag *(int)*: Page number for the first page of the article.
- last_page *(int)*: Page number for the last page of the article.

Currently the output prints as a regualar `data.frame`. You can make the output
prettier by using `tibble::as_tibble` or, for a single case, by using `tidyr::gather`:

```{r, results='asis'}
meta_data %>% knitr::kable()
```

### `find_authors`
Extracting the authors works in similar fashion:

```{r, results='asis'}
authors <- find_authors(jstor_example("sample_with_references.xml"))
knitr::kable(authors)
```

Here we have the following columns:

- *basename_id*: The same as above, used for matching articles.
- *prefix*: A prefix to the name.
- *given_name*: The given name of the author (i.e. `Albert` or `A.`).
- *surname*: The surname of the author (i.e. `Einstein`).
- *string_name*: Sometimes instead of given_name and surname, only a full string is
supplied, i.e.: `Albert Einstein`, or `Einstein, Albert`.
- *suffix*: A suffix to the name, as in `Albert Einstein, II.`.
- *author_number*: An integer representing the order of how the authors appeard in the data.

The number of rows matches the number of authors -- each author get its' own row.

### `find_references`
```{r}
references <- find_references(jstor_example("sample_with_references.xml"))
```

We have three columns:

- *basename_id*: Identifier, can be used for matching.
- *full_reference*: The content of each entry for reference.
- *author_names*: The names of the references authors, as far as specified by JSTOR.

Here I display the first 5 entries for each column:

```{r}
references %>% 
  purrr::map(head, 5)
```

This example shows several things: `basename_id` is identical among rows, since
it identifies the article and all references came from one article.

The content of references (`full_reference`) is in quite a raw state, quite often
the result of digitising scans via OCR. Very often, the first entry in this column
is something like "Bibliography" or "References", which is simply the heading 
within the article. In the above example there are several headings, because the
sample file doesn't follow a typical convention (it was published in 1922). 

The last column, `author_names`, contains strings with just the names of authors.
This data is generated by JSTOR/DfR and can be quite messy. In the above
example it only contains surnames, but no initials. Sometimes it might
contain only the first author, or none. Use at your own discretion.


### `find_footnotes`
```{r}
find_footnotes(jstor_example("sample_with_references.xml"))
```

Very commonly, articles either have footnotes or references. The sample file
used here does not have footnotes, therefore a simple `data.frame` with missing
footnotes is returned.

I will use another file to demonstrate footnotes.

```{r}
footnotes <- find_footnotes(jstor_example("sample_with_footnotes.xml"))
footnotes
```

As can be seen from the output above, for footnotes there is no extra column for
authors, only the id and the content of the footnotes.

### `find_fulltext`
The function to extract full texts can't be demonstrated with sample data, since
the full texts are only supplied upen special request with JSTOR. The function
guesses the encoding of the specified file via `readr::guess_encoding()`, reades
the whole file and returns a `data.frame` with `basename_id`, `full_text` and
`encoding`.

### Combining results
Different parts of meta-data can be combined by using `dplyr::left_join()`.

**Matching with authors**

```{r, results='asis'}
meta_data %>% 
  left_join(authors) %>%
  select(basename_id, article_title, pub_year, given_name, surname) %>% 
  knitr::kable()
```

**Matching with references**

```{r, results='asis'}
meta_data %>% 
  left_join(references) %>%
  select(basename_id, article_title, pub_year, author_names) %>% 
  head(10) %>% 
  knitr::kable()
```


