---
title: "Known Quirks of JSTOR Data"
author: "Thomas Klebel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Known Quirks of JSTOR Dat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Collecting all the quirks
Data from JSTOR/DfR is unlike most other data you encounter when doing text 
analysis. First and foremost, the data about the articles and books come from a
wide variety of journals or publishers. The level of detail and certain formats
vary because of this. `jstor` tries to deal with this situation with two
strategies:

- try to recognise the format and read data accordingly
- if this is not possible, read data as "raw" as possible, i.e. without any 
conversions

An example for the latter case are page numbers. Most of the time, the entries
for page numbers are simply `42`, or `61`. This is as expected, and could be
parsed as integers. Sometimes, there are characters present, like `M77`. This
would pose no problem either, we could simply extract all digits via regex and
parse as character. Unfortunately, sometimes the page is specified like this:
`v75i2p84`. Extracting all digits would result in `75284`, which is wrong by
a long shot. Since there might be other ways of specifying pages, `jstor` does
not attempt to parse the pages to integers when importing.

There are many other problems or pecularities like this. This vignette tries to
list as many as possible, and offer solutions for dealing with them.
Unfortunately I have neither the time nor the interest to wade through all the
data which you could get from DfR in order to find all possible quirks. The
following list is thus inevitably incomplete. If you encounter new 
quirks/pecularities, you are thus welcome to share them by sending me an email,
or [opening an issue at GitHub](https://github.com/tklebel/jstor/issues), and 
I will include them in future version of this vignette, so this vignette can
be a starting point for everybody who conducts text analysis with data from
JSTOR/DfR.


# Known quirks

```{r}
library(jstor)
```


## Page numbers
### Clean pages

### Total pages
As outlined above, page numbers come in very different forms. Besides this
problem, there is actually another issue. Imagine you would like to quantify 
the lenghts of articles. Obviously you will need information on the first
and the last page of the articles. Furthermore, the pages need to be
parsed properly: you will run into troubles if you calculate page numbers like
`75284 - 42 + 1`, in case the number was parsed badly. Parsing correctly is
unfortunately not enough. Things like "Errata" could come to haunt you.

There might be an article with `first_page = 42` and `last_page = 362`, which
would leave you puzzled as to if this can be true^[Although it sounds absurd,
this can acutally be true. There are some articles which are 200 pages long.
Obviously, they are not standard research articles. You will need to decide
if they fall into your sample or not]. There could be a simple explanation:
the article might start on page 42, and end on page 65, and there is furthermore
an erratum on page 362. Technically, `last_page = 362` is true then, but it
will cause problems for calculating the total number of pages. Quite often,
there is information in another column which could resolve this: `page_range`, 
which in this case would look like `42 - 65, 362`. 

A small helper to deal with those situations is `jst_get_total_pages()`. It
works for page ranges, but also for first and last pages:

```{r}
jst_get_total_pages(page_range = "42 - 65, 362")

jst_get_total_pages(first_page = 20, last_page = 39)

jst_get_total_pages(first_page = 20, last_page = NA)
```


## Unify journal id


## jst_augment
